### 2020-03-05 manuscript/supp-mat.tex
Added Scaling experiments, define function in supp. mat.

\documentclass{article}

\usepackage{url}

\usepackage{amssymb,amsfonts,amsmath,amsthm,mathtools}
\usepackage{adjustbox}
\usepackage{float}
\usepackage{caption}
\usepackage{mdframed}
\usepackage{lmodern}
\usepackage{bm,bbold}
\usepackage{xfrac, nicefrac}
\usepackage{xcolor}
\usepackage{lmodern}
\usepackage{enumitem}
\usepackage{pgfplots, pgf,tikz}
\usepackage[margin=60pt]{geometry}
\pdfinclusioncopyfonts=1

\definecolor{RED}{HTML}{EB6231}
\definecolor{YELLOW}{HTML}{E29D26}
\definecolor{BLUE}{HTML}{5D80B4}
\definecolor{LIGHTGREEN}{HTML}{6ABD9B}
\definecolor{GREEN}{HTML}{8FB03E}
\definecolor{PURPLE}{HTML}{BE1E2D}
\definecolor{BROWN}{HTML}{A97C50}
\definecolor{PINK}{HTML}{DA1C5C}
\captionsetup{width=0.85\textwidth}

\newcommand{\specialcell}[2][c]{%
	\begin{tabular}[#1]{@{}c@{}}#2\end{tabular}}

\DeclareMathOperator{\E}{\mathbb{E}}
\newcommand{\der}{\mathrm{d}}
\newcommand{\e}{\mathrm{e}}

% Time, effective population size and mutation rate.
\newcommand{\Ne}{N_{\mathrm{e}}}
\newcommand{\dnds}{\omega}
\newcommand{\Nsite}{\text{n}}
\newcommand{\Nstate}{\text{K}}

\newcommand{\x}{x}
\newcommand{\eq}{^{*}}
\newcommand{\dx}{\delta \x}

\begin{document}
\part*{Supplementary materials}

\section*{Genotype to phenotype map}
Define $\Nsite$ as the number of sites in the genotype sequence.
Each site can be in one of $\Nstate \geq 2$ states, where only $1$ state is defined the stable state, and $\Nstate - 1$ states are unstable.
For a given genotype sequence, define phenotype $0 \leq \x \leq 1$ as the current proportion of sites in the unstable state.
After a mutation, given that only one site can change at a time, the absolute change of $\x$ is either $0$ or $\dx=\sfrac{1}{\Nsite}$.
Define $\rho_{\x}(\dx)$ as the probability to get a change of phenotype equal to $\dx$, if the current phenotype is $\x$:\\
\begin{gather}
\begin{cases}
\dx &\text{ with probability } \rho_{\x}(\dx) = 1-x, \\
0 &\text{ with probability } \rho_{\x}(0) = \x \left[1 - \frac{1}{\Nstate - 1}\right], \\
-\dx &\text{ with probability } \rho_{\x}(-\dx) = \frac{x}{\Nstate - 1}.\\
\end{cases} \label{eq:proba}
\end{gather}

\section*{Selection coefficient}
$s(\x, \dx)$ is the selection coefficient of an effect $\dx$ if the current phenotype is $\x$:
\begin{align}
s(\x, \dx) &= \frac{f(\x + \dx) - f(\x)}{f( \x)}, \\
 &\simeq \frac{1}{f(\x)}\frac{\der f( \x)}{\der \x} \dx, \\
 &\simeq \frac{\der \ln (f(\x))}{\der \x} \dx, \label{eq:s_from_fitness}
\end{align}
where $f( \x)$ is the Wrightian fitness of phenotype $x$. And thus we also have:
\begin{gather}
s(\x, -\dx) \simeq - s(\x, \dx) \text{ from eq. \ref{eq:s_from_fitness}} \label{eq:s_minus_deltax}, \\
\iff S(\x, -\dx) \simeq - S(\x, \dx), \label{eq:S_minus_deltax}
\end{gather}
where $S(\x\eq, \dx) = 4\Ne s (\x\eq, \dx)$ is the scaled selection coefficient.

\section*{Probability of fixation}
The probability of fixation of a mutation with effect $\dx$, for a resident phenotype $\x$ is :
\begin{align}
p_{\text{fix}}(\x, \dx) &= \frac{ 2 s(\x, \dx)}{1 - \e^{-4\Ne s(\x, \dx)}}, \\
 &= \frac{ 2 s(\x, \dx)}{1 - \e^{-S(\x, \dx)}}. \label{eq:pfix}
\end{align}
And in the case of neutral mutations, the probability of fixation is:
\begin{gather}
p_{\text{fix}}(\x, 0) = \frac{ 1}{2 \Ne} \label{eq:pfix0}.
\end{gather}
And the ratio of probability of fixation between selected and neutral mutations is:
\begin{align}
\frac{p_{\text{fix}}(\x, \dx)}{p_{\text{fix}}(\x, 0)} & = \frac{ 2 \Ne 2 s(\x, \dx)}{1 - \e^{-S(\x, \dx)}} \text{ from eq. \ref{eq:pfix} and \ref{eq:pfix0}},\\
& = \frac{S(\x, \dx)}{1 - \e^{-S(\x, \dx)}} \label{eq:pfixratio}.
\end{align}
\section*{Equilibrium phenotype}
At equilibrium phenotype $\x\eq$, the expected selection coefficient of mutation that reached fixation must be $0$:
\begin{gather}
 0 = \E_{\dx} \left[ s(\x\eq, \dx) p_{\text{fix}}(\x\eq, \dx) \right], \\
\iff 0 = \frac{ 2 s(\x\eq, \dx)^2}{1 - \e^{-S(\x\eq, \dx)}}   \rho_{\x\eq}(\dx) + s(\x\eq, 0) \frac{ \rho_{\x\eq}(0)}{2\Ne} + \frac{ 2s(\x\eq, -\dx)^2}{1 - \e^{-S(\x\eq, -\dx)}} \rho_{\x\eq}(-\dx)\text{ from eq. \ref{eq:pfix} and \ref{eq:pfix0}}, \\
\Longrightarrow \frac{ 2s(\x\eq, \dx)^2}{1 - \e^{-S(\x\eq, \dx)}}   \rho_{\x\eq}(\dx) \simeq \frac{ - 2s(\x\eq, \dx)^2}{1 - \e^{S(\x\eq, \dx)}}   \rho_{\x\eq}(-\dx) \text{ from eq. \ref{eq:S_minus_deltax}}, \\
\iff \frac{ \rho_{\x\eq}(\dx)}{\rho_{\x\eq}(-\dx)} \simeq \e^{-S(\x\eq, \dx)} \frac{ \e^{-S(\x\eq, \dx)} - 1}{ \e^{-S(\x\eq, \dx)} \left( 1 - \e^{S(\x\eq, \dx)} \right)}, \\
\iff \ln \left( \frac{1 - \x\eq}{\x\eq} \right) + \ln (\Nstate-1) \simeq -S(\x\eq, \dx) \text{ from eq. \ref{eq:proba}} \label{eq:equilibrium}, \\
\iff \lambda_{K}(\x\eq) \simeq -S(\x\eq, \dx) \label{eq:equilibrium_lambda},
\end{gather}
where  $\lambda_{K}(\x\eq) = \ln \left( \frac{1 - \x\eq}{\x\eq} \right) + \ln (\Nstate-1)$.
\section*{Substitution rate ($\bm{\dnds}$) at equilibrium}
The substitution rate of selected mutations relative to neutral mutations  ($\dnds$) is by definition:
\begin{align}
\dnds & = \E_{\dx} \left[ \frac{p_{\text{fix}}(\x, \dx)}{p_{\text{fix}}(\x, 0)} \right], \\
 & = (1 - \x) \frac{ S(\x, \dx)}{1 - \e^{-S(\x, \dx)}} + \x \left(\frac{\Nstate - 2}{\Nstate - 1}\right) + \frac{\x}{\Nstate-1} \frac{ S(\x, -\dx)}{1 - \e^{-S(\x, -\dx)}} \text{ from eq. \ref{eq:proba}, \ref{eq:pfix} and \ref{eq:pfix0}}, \\
 & = (1 - \x) \frac{ S(\x, \dx)}{1 - \e^{-S(\x, \dx)}} - \frac{\x}{\Nstate-1}  \frac{ S(\x, \dx)}{1 - \e^{S(\x, \dx)}} +  \x \left(\frac{\Nstate - 2}{\Nstate - 1}\right) \text{ from eq. \ref{eq:S_minus_deltax}}.
\end{align}
$\dnds\eq$ at equilibrium is then determined by the phenotype at equilibrium $\x\eq$:
\begin{align}
\dnds\eq &= (1 - \x\eq) \frac{ S(\x\eq, \dx)}{1 - \e^{-S(\x\eq, \dx)}} - \frac{\x\eq}{\Nstate-1} \frac{ S(\x\eq, \dx)}{1 - \e^{S(\x\eq, \dx)}} + \x\eq \left(\frac{\Nstate - 2}{\Nstate - 1}\right) , \\
 &= \x\eq \left[ \frac{2 (\x\eq-1)  \lambda_{K}(\x\eq) }{\Nstate (\x\eq-1)+1} + \frac{\Nstate - 2}{\Nstate - 1}\right] \text{ from eq. \ref{eq:equilibrium}}.
\end{align}
\begin{center}
	\begin{tikzpicture}[
	declare function={
		omega(\x,\k)= \x*(2*(\x-1)*(ln(\k-1)+ln((1-\x)/\x))/(\k*(\x-1)+1)+ (\k-2)/(\k-1));
	},]
	\begin{axis}[
	ylabel={$\dnds\eq$},
	xlabel={$\x\eq$},
	domain=0:1.0,
	ymin=0.0, ymax=1.1,
	samples=100,
	legend entries={$\Nstate=2$, $\Nstate=4$, $\Nstate=8$, $\Nstate=16$},
	legend cell align=left,
	minor tick num=2,
	axis x line=bottom,
	axis y line=left,
	legend style={at={(0.6,0.34)},anchor=north west}
	]
	\addplot[line width=2.0pt, color=BLUE]{omega(x, 2)};
	\addplot[line width=2.0pt, color=YELLOW]{omega(x, 4)};
	\addplot[line width=2.0pt, color=GREEN]{omega(x, 8)};
	\addplot[line width=2.0pt, color=RED]{omega(x, 16)};
	\end{axis}
	\end{tikzpicture}
\end{center}
And the derivative of $\dnds\eq$ w.r.t to $\x\eq$ is: 
\begin{gather}
\frac{\der \dnds\eq}{\der \x\eq} = 2 \left[ \frac{\Nstate (\x\eq- 1) + 1+\left[\Nstate (\x\eq-1)^2+2 \x\eq-1\right] \lambda_{K}(\x\eq) }{(\Nstate (\x\eq-1)+1)^2}\right] + \frac{\Nstate - 2}{\Nstate - 1} \label{eq:dw_dx}.
\end{gather}
\begin{center}
	\begin{tikzpicture}[
	declare function={
		domega(\x,\k)= 2*(\k*(\x-1)+1 + (\k*(\x-1)^2+2*\x-1)*(ln(\k-1) + ln((1-\x)/\x)))/(\k*(\x - 1) + 1)^2 + (\k-2) / (\k-1);
	},]
	\begin{axis}[
	ylabel={$\frac{\der \dnds\eq}{\der \x\eq}$},
	xlabel={$\x\eq$},
	domain=0.0:1.0,
	ymin=0, ymax=6,
	samples=50,
	legend entries={$\Nstate=2$, $\Nstate=4$, $\Nstate=8$, $\Nstate=16$},
	legend cell align=left,
	minor tick num=2,
	axis x line=bottom,
	axis y line=left,
	legend style={at={(0.5,0.95)},anchor=north west}
	]
	\addplot[line width=2.0pt, color=BLUE]{domega(x, 2)};
	\addplot[line width=2.0pt, color=YELLOW]{domega(x, 4)};
	\addplot[line width=2.0pt, color=GREEN]{domega(x, 8)};
	\addplot[line width=2.0pt, color=RED]{domega(x, 16)};
	\end{axis}
	\end{tikzpicture}
\end{center}

\section*{Response of equilibrium after a change in $\bm{\Ne}$}
Define the function $G(\x, \Ne )$ as:
\begin{gather}
G(\x, \Ne ) \equiv \ln \left( \frac{1 - \x}{\x} \right) + \ln (\Nstate-1) + 4 \Ne s(\x, \dx), 
\end{gather}
The equilibrium equation (eq. \ref{eq:equilibrium}) states that $G(\x\eq, \Ne )=0$, meaning that $\x\eq$ is implicitly a function of $\Ne$:
\begin{gather}
G(\x\eq(\Ne), \Ne ) = 0, \\
\Longrightarrow \frac{\partial G(\x\eq, \Ne ) }{\partial \x\eq }\frac{ \der \x\eq}{\der \Ne} + \frac{ \partial G(\x\eq, \Ne )}{\partial \Ne} = 0, \\
\iff \left[ \frac{\partial \ln (1-\x\eq) }{\partial \x\eq } - \frac{\partial \ln (\x\eq) }{\partial \x\eq}  + 4 \Ne \frac{ \partial s(\x\eq, \dx) }{\partial \x\eq } \right]\frac{ \der \x\eq}{\der \Ne} + 4 s(\x\eq, \dx) = 0, \\
\iff \left[ \frac{1}{(1 - \x\eq) \x\eq} + 4 \Ne \frac{ \partial^2 \ln f(\x\eq) }{\partial {\x\eq}^2}\dx \right]\frac{ \der \x\eq}{\der \Ne}  = - 4 \frac{ \partial \ln f(\x\eq) }{\partial {\x\eq}}\dx \text{ from eq. \ref{eq:s_from_fitness}}, \\
\iff 4 \dx \left[ \frac{1}{4 \dx \Ne  (1 - \x\eq) \x\eq} + \frac{ \partial^2 \ln f(\x\eq) }{\partial {\x\eq}^2} \right] \Ne \frac{ \der \x\eq}{\der \Ne}  = - 4 \dx \frac{ \partial \ln f(\x\eq) }{\partial {\x\eq}}, \\
\iff \frac{ \der \x\eq}{\der \ln (\Ne)}  = - \frac{\frac{ \partial \ln f(\x\eq) }{\partial {\x\eq}}}{\frac{1}{4 \dx \Ne  (1 - \x\eq) \x\eq} + \frac{ \partial^2 \ln f(\x\eq) }{\partial {\x\eq}^2}}  \label{eq:dx_dlnNe}.
\end{gather}
Giving the equation for the response of phenotype at equilibrium after a change of effective population size.
Together, the response of substitution rate at equilibrium, after a change of effective population size can be obtain as:
\begin{align}
\frac{\der \dnds\eq}{\der \ln (\Ne)} & = \frac{\der \dnds\eq}{\der \x\eq} \frac{ \der \x\eq}{\der \ln (\Ne)}, \\
 &= - \frac{\der \dnds\eq}{\der \x\eq} \frac{\frac{ \partial \ln f(\x\eq) }{\partial {\x\eq}}}{\frac{1}{4 \dx \Ne  (1 - \x\eq) \x\eq} + \frac{ \partial^2 \ln f(\x\eq) }{\partial {\x\eq}^2}} \text{ from eq. \ref{eq:dx_dlnNe}} \label{eq:dw_dlnNe}.
\end{align}
\section*{Phenotype to fitness mapping}
From a given phenotype $\x$, define the fitness as :
\begin{gather}
 f(\x) = \frac{1}{1 + \e^{\beta(\alpha + \kappa \x)}}, \label{eq:fitness}
\end{gather}
where $\alpha = \Delta G_0 < 0$ is the difference in free energy between stable and unstable protein when all sites are stable. $\kappa$ is the expected change in $\Delta G$ when all sites are unstable. By definition, $\kappa \dx = \sfrac{\kappa}{\Nsite} = \Delta \Delta G $\\
The derivative of fitness w.r.t to phenotype is:
\begin{align}
 \frac{\partial \ln (f(\x))}{\partial \x}  & = - \frac{\partial \ln \left( 1 + \e^{\beta(\alpha + \kappa \x)} \right)}{\partial \x} \text{ from eq. \ref{eq:fitness}}, \\
 & = - \beta \kappa \frac{\e^{\beta(\alpha + \kappa \x)}}{1 + \e^{\beta(\alpha + \kappa \x)}} \label{eq:df_dx}.
\end{align}
And the second derivative of fitness w.r.t to phenotype is:
\begin{align}
\frac{\partial^2 \ln (f(\x))}{\partial {\x}^2} & = - \beta \kappa \frac{\partial}{\partial \x} \left( \frac{\e^{\beta(\alpha + \kappa \x)}}{1 + \e^{\beta(\alpha + \kappa \x)}} \right) \text{ from eq. \ref{eq:df_dx}}, \\
 & = - \beta \kappa  \beta \kappa \frac{\e^{\beta(\alpha + \kappa \x)}}{\left( 1 + \e^{\beta(\alpha + \kappa \x)}\right)^2}, \\
 & = \frac{\beta \kappa}{1 + \e^{\beta(\alpha + \kappa \x)}} \frac{ \partial \ln f(\x) }{\partial {\x}} \text{ from eq. \ref{eq:df_dx}} \label{eq:df2_dx2}.
\end{align}
The equilibrium phenotype ($\x\eq$) is :
\begin{gather}
\lambda_{K}(\x\eq) = 4\Ne \beta \kappa \frac{\e^{\beta(\alpha + \kappa \x\eq)}}{1 + \e^{\beta(\alpha + \kappa \x\eq)}}  \dx   \text{ from eq. \ref{eq:equilibrium_lambda} and \ref{eq:df_dx}}.
\end{gather}
Using $\Ne=10^4$, $\beta=1.686$, $\alpha = -118$, $\Nsite=300$, $\kappa=300$, we have the following:
\begin{center}
\begin{tikzpicture}[
	declare function={
		lambda(\x,\k)= ln(\k-1) + ln((1-\x)/\x);
	},]
	\begin{axis}[
	xlabel={$\x$},
	ymin=0.0, ymax=5.0,
	samples=100,
	legend entries={
		$4\Ne \beta \kappa \frac{\e^{\beta(\alpha + \kappa \x)}}{1 + \e^{\beta(\alpha + \kappa \x)}} \dx$,
		$\lambda_{2}(\x)$,
		$\lambda_{4}(\x)$,
		$\lambda_{8}(\x)$,
		$\lambda_{16}(\x)$},
	legend cell align=left,
	minor tick num=2,
	axis x line=bottom,
	axis y line=left,
	legend style={at={(0.5,0.95)},anchor=north west}
	]
	\addplot[domain=0.0:0.38, line width=2.0pt, color=BLUE]{ 4 * 1000 * 1.686 *exp(1.686 * (-118 + 300 *x))};
	\addplot[domain=0.0:0.5,line width=2.0pt, color=YELLOW]{ lambda(x,2) };
	\addplot[domain=0.0:0.8,line width=2.0pt, color=GREEN]{ lambda(x,4) };
	\addplot[domain=0.0:0.9,line width=2.0pt, color=RED]{ lambda(x,8) };
	\addplot[domain=0.0:1.0,line width=2.0pt, color=LIGHTGREEN]{ lambda(x,16) };
	\end{axis}
\end{tikzpicture}
\end{center}
\section*{Approximations}
We assume that $\Ne \gg 1$, $\alpha + \kappa \gg 1$ and $\kappa \dx \simeq 1$.
\begin{gather}
-S(\x\eq, \dx) \simeq \lambda_{\Nstate}(\x\eq)\text{ from eq. \ref{eq:equilibrium_lambda}}, \\
\iff s(\x\eq, \dx) \simeq -\frac{\lambda_{\Nstate}(\x\eq)}{4 \Ne}, \\
\iff \frac{\der \ln (f(\x))}{\der \x} \simeq - \frac{\lambda_{\Nstate}(\x\eq)}{4 \Ne \dx}\text{ from eq. \ref{eq:s_from_fitness}}, \label{eq:approx_df_dx}\\
\iff \beta \kappa \frac{\e^{\beta(\alpha + \kappa \x)}}{1 + \e^{\beta(\alpha + \kappa \x)} } \simeq \frac{\lambda_{\Nstate}(\x\eq)}{4 \Ne \dx}\text{ from eq. \ref{eq:df_dx}}, \\
\iff \frac{\e^{\beta(\alpha + \kappa \x)}}{1 + \e^{\beta(\alpha + \kappa \x)}} \simeq \frac{\lambda_{\Nstate}(\x\eq)}{4 \Ne \dx \beta \kappa}, \\
\Longrightarrow \frac{\e^{\beta(\alpha + \kappa \x)}}{1 + \e^{\beta(\alpha + \kappa \x)}} \ll 1\text{ given that $\Ne \gg 1$ and $\kappa \dx \simeq 1$}, \\
\Longrightarrow \e^{\beta(\alpha + \kappa \x)} \ll 1. \label{eq:approx_exp}
\end{gather}
Moreover, we can get the following:
\begin{gather}
\frac{\der \ln (f(\x))}{\der \x} \simeq - \frac{\lambda_{\Nstate}(\x\eq)}{4 \Ne \dx}\text{ from eq. \ref{eq:approx_df_dx}}, \\
\iff \frac{ \partial^2 \ln f(\x) }{\partial {\x}^2} \simeq - \frac{\lambda_{\Nstate}(\x\eq)\left( 1 + \e^{\beta(\alpha + \kappa \x)} \right)}{4 \Ne \dx \beta \kappa }\text{ from eq. \ref{eq:df2_dx2}}, \\
\Longrightarrow \frac{ \partial^2 \ln f(\x) }{\partial {\x}^2} \simeq - \frac{\lambda_{\Nstate}(\x\eq)}{4 \Ne \dx \beta \kappa }\text{ from eq. \ref{eq:approx_exp}}, \\
\iff \frac{ \partial^2 \ln f(\x) }{\partial {\x}^2} \simeq - \lambda_{\Nstate}(\x\eq) \frac{(1 - \x\eq)\x\eq }{\beta \kappa} \frac{1}{4 \dx \Ne  (1 - \x\eq) \x\eq}, \\
\Longrightarrow \frac{ \partial^2 \ln f(\x) }{\partial {\x}^2} \simeq - \lambda_{\Nstate}(\x\eq) \dx \frac{(1 - \x\eq)\x\eq }{\beta } \frac{1}{4 \dx \Ne  (1 - \x\eq) \x\eq}\text{ given that $\kappa \dx \simeq 1$}, \\
\Longrightarrow \left| \frac{ \partial^2 \ln f(\x) }{\partial {\x}^2} \right| \gg \left| \frac{1}{4 \dx \Ne  (1 - \x\eq) \x\eq} \right|\text{ given that $\dx \ll 1$}. \label{eq:approx_df2_df} \\
\end{gather}
Together, these approximations leads to the following response in equilibrium phenotype at change in $\Ne$ as:
\begin{align}
\frac{ \der \x\eq}{\der \ln (\Ne)} & = - \frac{\frac{ \partial \ln f(\x\eq) }{\partial {\x\eq}}}{\frac{1}{4 \dx \Ne  (1 - \x\eq) \x\eq} + \frac{ \partial^2 \ln f(\x\eq) }{\partial {\x\eq}^2}}\text{ from eq. \ref{eq:dw_dlnNe}}, \\
& \simeq - \frac{\frac{ \partial \ln f(\x\eq) }{\partial {\x\eq}}}{\frac{ \partial^2 \ln f(\x\eq) }{\partial {\x\eq}^2}}\text{ from eq. \ref{eq:approx_df2_df}}, \\
& \simeq - \frac{1 + \e^{\beta(\alpha + \kappa \x\eq)}}{\beta \kappa}\text{ from eq. \ref{eq:df2_dx2}}, \\
& \simeq - \frac{1}{\beta \kappa}\text{ from eq. \ref{eq:approx_exp}}. \label{eq:approx_dx_dlnNe}
\end{align}
Moreover, given that the number of state if large enough $\Nstate \gg 1$, the response in equilibrium $\dnds$ due to change in phenotype can be approximated as:
\begin{align}
\frac{ \der \dnds\eq}{\der \x\eq}  & = 2 \left[ \frac{\Nstate (\x\eq- 1) + 1+\left[\Nstate (\x\eq-1)^2+2 \x\eq-1\right] \lambda_{K}(\x\eq) }{(\Nstate (\x\eq-1)+1)^2}\right] + \frac{\Nstate - 2}{\Nstate - 1}\text{ from eq. \ref{eq:approx_exp}}, \\
& \simeq \frac{2 \lambda_{K}(\x\eq)}{\Nstate} + 1, \\
& \simeq 1. \label{eq:approx_domega_dx}
\end{align}
Finally,
\begin{align}
\frac{ \der \dnds\eq}{\der \ln (\Ne)}  & = \frac{ \der \dnds\eq}{\der \x\eq}  \frac{ \der \x\eq}{\der \ln (\Ne)}, \\
  & \simeq - \frac{1}{\beta \kappa} \text{from eq. \ref{eq:approx_dx_dlnNe} and \ref{eq:approx_domega_dx}}.
\end{align}
\section*{Numerical results}
\begin{center}
\begin{tikzpicture}
	\begin{axis}[
	xlabel={$\x$},
	ymin=0.0, ymax=1.0,
	samples=100,
	legend entries={$\ln(\frac{1 - \x}{\x})$,
		$\alpha = -118\text{, }\kappa = 300\text{, }\Ne = 10^{4}$,
		$\alpha = -118\text{, }\kappa = 300\text{, }\Ne = 10^{5}$,
		$\alpha = -16.7\text{, }\kappa = 30\text{, }\Ne = 10^{4}$,
		$\alpha = -16.7\text{, }\kappa = 30\text{, }\Ne = 10^{5}$},
	legend cell align=left,
	minor tick num=2,
	axis x line=bottom,
	axis y line=left,
	legend style={at={(0.5,0.95)},anchor=north west}
	]
	\addplot[domain=0.3:0.6,line width=2.0pt, color=BLUE]{ ln((1-x)/x) };		\addplot[domain=0.3:0.38, line width=2.0pt, color=RED]{4*10000*1.686*exp(1.686*(-118+300*x))};
	\addplot[dashed,domain=0.3:0.37, line width=2.0pt, color=RED]{4*100000*1.686*exp(1.686*(-118+300*x))};
	\addplot[domain=0.3:0.39, line width=2.0pt, color=GREEN]{4*10000*0.1*1.686*exp(1.686*(-16.71+30*x))};
	\addplot[dashed,domain=0.3:0.35, line width=2.0pt, color=GREEN]{4*100000*0.1*1.686*exp(1.686*(-16.71+30*x))};
	\end{axis}
\end{tikzpicture}
\end{center}
\end{document}