### 2020-05-09 manuscript/figures.tex
Added Figures and updated manuscript

\documentclass{beamer}

\usepackage{amssymb,amsfonts,amsmath,amsthm,mathtools}
\usepackage{lmodern}
\usepackage{xfrac, nicefrac}
\usepackage{pgfplots, pgf,tikz}
\usepgfplotslibrary{fillbetween}
\usebackgroundtemplate{\tikz\node[opacity=0]{};}
\setbeamertemplate{footline}[frame number]{}
\setbeamertemplate{navigation symbols}{}
\setbeamertemplate{footline}{}
\usefonttheme{serif}

\include{notations}
\pgfplotsset{every axis/.append style={line width=1pt}}
\pgfplotscreateplotcyclelist{colors}{LIGHTGREEN\\YELLOW\\RED\\GREEN\\BLUE\\}

\begin{document}
	\begin{frame}
	\centering
	\begin{tikzpicture}
	\begin{axis}[
	width=\textwidth,
	height=\axisdefaultheight,
	xlabel={$x$},
	ymin=0.0, ymax=5.0,
	samples=200,
	legend entries={$\ln(\frac{1 - x}{x}) + \ln(19)$,
		${\alpha = -118, \gamma = 1,\Ne = 10^{4}}$,
		${\alpha = -118, \gamma = 1,\Ne = 10^{5}}$,
		${\alpha = -15.7, \gamma = 0.1,\Ne = 10^{4}}$,
		$\tiny{\alpha = -15.7, \gamma = 0.1,\Ne = 10^{5}}$,
	},
	legend cell align=left,
	minor tick num=2,
	axis x line=bottom,
	axis y line=left,
	legend style={at={(0.05,0.25)},anchor=south west}
	]
	\addplot[domain=0.1:0.39,line width=2.0pt, color=BLUE]{ ln(19*(1-x)/x) };
	\addplot[name path=A, domain=0.1:0.38, line width=2.0pt, color=RED]{4*10000*1.686*exp(1.686*(-118+300*x))};
	\addplot[name path=B, dashed,domain=0.3:0.37, line width=2.0pt, color=RED]{4*100000*1.686*exp(1.686*(-118+300*x))};
	\addplot[name path=C,domain=0.1:0.39, line width=2.0pt, color=GREEN]{4*10000*0.1*1.686*exp(1.686*(-15.71+30*x))};
	\addplot[name path=D,dashed,domain=0.1:0.35, line width=2.0pt, color=GREEN]{4*100000*0.1*1.686*exp(1.686*(-15.71+30*x))};
	\addplot[fill=RED, opacity=0.2] fill between[ of = A and B];
	\addplot[fill=GREEN, opacity=0.2] fill between[ of = C and D];
	\end{axis}
	\end{tikzpicture}
\end{frame}

\begin{frame}
\centering
\begin{tikzpicture}[
declare function={
	omega(\x,\k)= \x*(2*(\x-1)*(ln(\k-1)+ln((1-\x)/\x))/(\k*(\x-1)+1)+ (\k-2)/(\k-1));
},]
\begin{axis}[
width=\textwidth,
height=\axisdefaultheight,
ylabel={$\omega\eq$},
xlabel={$x\eq$},
domain=0:1.0,
ymin=0.0, ymax=1.1,
samples=200,
legend entries={$\Nstate=2$, $\Nstate=5$, $\Nstate=10$, $\Nstate=20$},
legend cell align=left,
minor tick num=2,
axis x line=bottom,
axis y line=left,
legend style={at={(0.5,0.4)},anchor=north west}
]
\addplot[line width=2.0pt, color=BLUE]{omega(x, 2)};
\addplot[line width=2.0pt, color=YELLOW]{omega(x, 5)};
\addplot[line width=2.0pt, color=GREEN]{omega(x, 10)};
\addplot[line width=2.0pt, color=RED]{omega(x, 20)};
\end{axis}
\end{tikzpicture}
\end{frame}

\begin{frame}
\centering
\begin{tikzpicture}[
declare function={
domega(\x,\k)= 2*(\k*(\x-1)+1 + (\k*(\x-1)^2+2*\x-1)*(ln(\k-1) + ln((1-\x)/\x)))/(\k*(\x - 1) + 1)^2 + (\k-2) / (\k-1);
},]
\begin{axis}[
width=\textwidth,
height=\axisdefaultheight,
ylabel={$\frac{\der \omega\eq}{\der x\eq}$},
xlabel={$x\eq$},
domain=0.0:1.0,
ymin=0, ymax=6,
samples=200,
legend entries={$\Nstate=2$, $\Nstate=5$, $\Nstate=10$, $\Nstate=20$},
legend cell align=left,
minor tick num=2,
axis x line=bottom,
axis y line=left,
legend style={at={(0.5,0.8)},anchor=north west}
]
\addplot[line width=2.0pt, color=BLUE]{domega(x, 2)};
\addplot[line width=2.0pt, color=YELLOW]{domega(x, 5)};
\addplot[line width=2.0pt, color=GREEN]{domega(x, 10)};
\addplot[line width=2.0pt, color=RED]{domega(x, 20)};
\end{axis}
\end{tikzpicture}
\end{frame}

\begin{frame}
\centering
\begin{tikzpicture}[
declare function={
lambda(\x,\k)= ln(\k-1) + ln((1-\x)/\x);
}]
\begin{axis}[
width=\textwidth,
height=\axisdefaultheight,
xlabel={$x$},
ymin=0.0, ymax=5.0,
samples=200,
legend entries={
$4\Ne \beta \gamma \frac{\e^{\beta(\alpha + \Nsite \gamma x)}}{1 + \e^{\beta(\alpha + \Nsite \gamma x)}}$,
$\lambda_{2}(x)$,
$\lambda_{5}(x)$,
$\lambda_{10}(x)$,
$\lambda_{20}(x)$},
legend cell align=left,
minor tick num=2,
axis x line=bottom,
axis y line=left,
legend style={at={(0.7,0.95)},anchor=north west}
]
\addplot[domain=0.0:0.38, line width=2.0pt, color=BLUE]{ 4 * 1000 * 1.686 *exp(1.686 * (-118 + 300 *x))};
\addplot[domain=0.0:0.5,line width=2.0pt, color=YELLOW]{ lambda(x,2) };
\addplot[domain=0.0:0.8,line width=2.0pt, color=GREEN]{ lambda(x,5) };
\addplot[domain=0.0:0.9,line width=2.0pt, color=RED]{ lambda(x,10) };
\addplot[domain=0.0:1.0,line width=2.0pt, color=LIGHTGREEN]{ lambda(x,20) };
\end{axis}
\end{tikzpicture}
\end{frame}

\begin{frame}
\begin{figure}[H]
\centering
\begin{tikzpicture}[scale = 1.0]
\begin{axis}[
width=\textwidth,
height=\axisdefaultheight,
ylabel={Fitness $f(x)$},
xlabel={Phenotype $x$},
cycle list name=colors,
domain=0:1,
samples=200,
legend cell align=left,
minor tick num=2,
axis x line=bottom,
axis y line=left,
legend entries={$f(x) = \frac{1}{1 + \e^{\beta(\alpha + \Nsite \gamma x)}}$},
legend style={at={(0.4,0.5)},anchor=south west}
]
\addplot[line width=2.0pt, color=BLUE]{ 1 / (1+exp(1.686*(-12 + 30 * x)))};
\addplot[black]{1.0};
\end{axis}
\end{tikzpicture}
\end{figure}
\end{frame}

\end{document}